Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-03-24 2_SSL_EES/article/article_supplementary.tex
final commit????

\begin{table}[]
\small
    \begin{tabularx}{\textwidth}{ |X|X|l|l|l|l|l|l|X|}
        \hline
            \textit{Parameter} & \textit{Unit} & \textit{2003} & $\pm [\%]$ & \textit{2012} & $\pm [\%]$ & \textit{2020} & $\pm [\%]$ & Source \\
        \hline
            Cleanroom Cost & \text{USD}/m$^2$ & 3000 & +16,-16 & 3000 & +16,-16 & 3000 & +16,-16 & \cite{mddi1997cleanroom}\cite{ledcomv2} \newline \cite{bakshi2009euv}\cite{gajera2006process} \\
        \hline
            Person-hour & FTE & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & I \\
        \hline
            Equip. Discount & \% of \text{USD} & 0\% & +25,-25 & 0\% & +25,-25 & 0\% & +25,-25 & I \newline \cite{Appleyard_2001} \\
        \hline
            Overall Yield & \% & 100\% & +25,-25 & 100\% & +25,-15 & 100\% & +25,-25 & I, \cite{lumi2012yield}\cite{ledsmag2012} \newline \cite{systemplus2015reverse}\cite{ledcomv2} \\
        \hline
            Inspec. Yield Savings & \%/inspec. & 0.5\% & +80,-80 & 0.5\% & +80,-80 & 0.5\% & +80,-80 & \cite{mckinseyyield} \\
        \hline
            Overall Throughput & UPH or h$^{-1}$ & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & Datasheets \\
        \hline
            Wafer Diameter & mm & 100 & +0,-50 & 150 & +33.3,-33.3 & 200 & +0,-25 & I and \cref{fig:wafers} \\
        \hline
            Edge Exclusion & mm & 7 & +0,-50 & 5 & +40,-0 & 5 & +40,-20 & \cite{ledsmagexclusion}\cite{rubiconexclusion} \newline \cite{xiamenexclusion}\cite{american2007annual} \\
        \hline
            Cutting Width & $\mu$m & 100 & +50,-25 & 75 & +33.3,-20 & 20 & +300,-50 & \cite{masaki2000division}\cite{ils2005width} \newline \cite{photonics2010width}\cite{discowidth} \\
        \hline
            Metal Prices & \text{USD}/kg & 100\% & +50,-0 & 100\% & +50,-0 & 100\% & +50,-0 & Datasheets \\
        \hline
            Electricity Price & \text{USD}/kWh & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & \cite{eia2000electric}\cite{eia2019electric} \\
        \hline
            Saph. Subst. Price & \text{USD} & 40 & +12.5,-12.5 & 10 & +100,-20 & 3 & +66.6,-33.3 & \cref{fig:wafers} \\
        \hline
            Phosphor Prices & \text{USD}/g & 100\% & +50,-50 & 150 & +0,-0 & 200 & +0,-0 & I, \cite{yole_phosphor_2012}\cite{yole2017phosphor} \\
        \hline
        \end{tabularx}
    \caption{Cost model sensitivity analysis parameter list. The results of the sensitivity analysis for the parameters in this table are presented in \cref{fig:sensitivity}. The units for the values in columns \textit{2003}-\textit{2020} are indicated in the column \textit{Units}. If values in columns \textit{2003}-\textit{2020} are instead given in \%, this indicates that the parameters were varied by a set percentage from their respective model baselines. Abbreviations: FTE - full-time equivalent; UPH - units per hour; Equip. Discount - equipment discount (sales rebate for large purchases); Inspec. Yield Savings - yield savings from inspection (early detection and alleviation of issues in the manufacturing workflow).}
    \label{tab:sensitivity}
\end{table}

@@

\begin{table}[H]
    \caption{Banana}
    \begin{tabularx}{\textwidth}{ |l|X|X|l|l| }
        \multicolumn{5}{c}{Forward Voltage Efficiency (VFE) $\eta_{Vf}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of Improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1999 & Indium Tin Oxide \newline Current spreading layer & Contact resistance & Yes & \cite{margalith1999indium}\\
        \hline
            1998 & Digitated electrodes & Contact resistance & No & \cite{steigerwald2001electrode} \\
        \hline
            Ongoing & Epitaxy improvements \newline and better doping & Polarization and \newline bulk resistance & No & I \\
        \hline
            1998 & Silver p-Contacts & Contact resistance & No & \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Light extraction efficiency (LEE) $\eta_{LE}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            $<$ 2003 & Optimization for \newline cavity effects & Reduces self-interference of quantum well & No & I, \cite{Shen2003} \\
        \hline
            1993 & Chip surface randomization & Total reflection and absorption & No & I, \cite{bergh1973surface}\cite{Schnitzer1993} \\
        \hline
            1993 & Thin-film chip architecture & Absorption & No & I, \cite{Schnitzer1993} \\
        \hline
            1996 & Patterned sapphire substrate  & Total reflection and absorption & Yes & I, \cite{Tadatomo2001} \\
        \hline
            Ongoing & Chip design for high LEE & Total reflection and absorption & No & I, \cite{Haerle2004} \\
        \hline
            $\sim$2000 & Silver p-contacts & Absorption & No & I, \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Internal quantum efficiency (IQE) $\eta_{IQ}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1994 & Double heterostructure & Higher Radiative recombination probability & No & \cite{Nakamura1994} \\
        \hline
            1996 & Multiple quantum well & Higher Radiative recombination probability & No & \cite{Koike1996} \\
        \hline
            Ongoing & Active region \newline Doping & Radiative recombination probability & No & I, \cite{schubert2018light} \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Radiative recombination probability & No & I \\
        \hline
            Ongoing & Chip architecture \newline Improvements & Radiative recombination probability & No & I \\
        \hline
        \multicolumn{5}{c}{Light Conversion efficiency (CE) $\eta_{C}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            Ongoing & Lower current density & Current density & No & I \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Charge distribution \newline Defect density & No & I, \cite{bhardwaj2016progress} \\
        \hline
            Ongoing & Chip architecture \newline improvements  & Charge distribution \newline Defect density & No & I, \cite{Wildeson2017} \\
        \hline
            $<$ 2017 & Defect getting \newline underlayer & Defect Density & No & I, \cite{haller2017burying} \\
        \hline
        \end{tabularx}
    LED innovations and technology improvements affecting key white LED device sub-efficiencies. The table includes innovations and technology improvements affecting forward voltage efficiency, light extraction efficiency, internal quantum efficiency, and light conversion efficiency. For a list of innovations affecting spectral efficiency, see Table 1 in the main article. \textit{Year} column indicates the first instance of application of corresponding invention in white LEDs. 'Ongoing' indicates improvements that are incremental in nature and have been ongoing since the earliest days of LED manufacturing, with no individual breakthroughs identified. \textit{Spillover} column indicates if the innovation involved technology spillovers. \textit{Source} column indicates the source of information about the innovation or improvement, with 'I' indicating expert interviews as such a source.
    \label{tab:innovations}
\end{table}

@@

\begin{table}[h!]
\small
    \centering
    \begin{tabular}{|l|l|l|l|l|}
    \hline
        \textit{\#} & \textit{Sector} & \textit{Role} & \textit{Country} & \textit{Expertise} \\ \hline
        1 & Academia & Sr. researcher & UK & Epitaxy \\ \hline
        2 & Industry & Consultant, former sr. researcher & USA & Device architecture \\ \hline
        3 & Industry & Consultant, former head of R\&D & Germany & Epitaxy \\ \hline
        4 & Academia & Professor & Austria & Phosphors \\ \hline
        5 & Industry & Consultant, former head of R\&D & USA & Device architecture \\ \hline
        6 & consulting & Consultant, former sr. technical advisor & USA & Device architecture \\ \hline
        7 & Academia & Professor & Germany & Phosphors \\ \hline
        8 & government & R\&D manager & USA & Device architecture \\ \hline
        9 & consulting & Consultant & USA & Device applications \\ \hline
        10 & Academia & Professor & France & Device physics \\ \hline
        11 & Industry & Sr scientist, former head of R\&D & USA & Device architecture \\ \hline
        12 & Industry & Principal scientist & Germany & Phosphors \\ \hline
        13 & Industry & Former head of R\&D & USA & Phosphors \\ \hline
    \end{tabular}
    \label{tab:interviews}
    \caption{Anonymized list of LED experts interviewed for this study. Abbreviations: sr. - senior}
\end{table}

>>>>

\begin{table}[]
\small
    \caption{Cost model sensitivity analysis parameter list. The results of the sensitivity analysis for the parameters in this table are presented in \cref{fig:sensitivity}.}
    \vspace{5mm}
    \begin{tabularx}{\textwidth}{ |X|X|l|l|l|l|l|l|X|}
        \hline
            \textit{Parameter} & \textit{Unit} & \textit{2003} & $\pm [\%]$ & \textit{2012} & $\pm [\%]$ & \textit{2020} & $\pm [\%]$ & Source \\
        \hline
            Cleanroom Cost & \text{USD}/m$^2$ & 3000 & +16,-16 & 3000 & +16,-16 & 3000 & +16,-16 & \cite{mddi1997cleanroom}\cite{ledcomv2} \newline \cite{bakshi2009euv}\cite{gajera2006process} \\
        \hline
            Person-hour & FTE & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & I \\
        \hline
            Equip. Discount & \% of \text{USD} & 0\% & +25,-25 & 0\% & +25,-25 & 0\% & +25,-25 & I \newline \cite{Appleyard_2001} \\
        \hline
            Overall Yield & \% & 100\% & +25,-25 & 100\% & +25,-15 & 100\% & +25,-25 & I, \cite{lumi2012yield}\cite{ledsmag2012} \newline \cite{systemplus2015reverse}\cite{ledcomv2} \\
        \hline
            Inspec. Yield Savings & \%/inspec. & 0.5\% & +80,-80 & 0.5\% & +80,-80 & 0.5\% & +80,-80 & \cite{mckinseyyield} \\
        \hline
            Overall Throughput & UPH or h$^{-1}$ & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & Datasheets \\
        \hline
            Wafer Diameter & mm & 100 & +0,-50 & 150 & +33.3,-33.3 & 200 & +0,-25 & I and \cref{fig:wafers} \\
        \hline
            Edge Exclusion & mm & 7 & +0,-50 & 5 & +40,-0 & 5 & +40,-20 & \cite{ledsmagexclusion}\cite{rubiconexclusion} \newline \cite{xiamenexclusion}\cite{american2007annual} \\
        \hline
            Cutting Width & $\mu$m & 100 & +50,-25 & 75 & +33.3,-20 & 20 & +300,-50 & \cite{masaki2000division}\cite{ils2005width} \newline \cite{photonics2010width}\cite{discowidth} \\
        \hline
            Metal Prices & \text{USD}/kg & 100\% & +50,-0 & 100\% & +50,-0 & 100\% & +50,-0 & Datasheets \\
        \hline
            Electricity Price & \text{USD}/kWh & 100\% & +50,-50 & 100\% & +50,-50 & 100\% & +50,-50 & \cite{eia2000electric}\cite{eia2019electric} \\
        \hline
            Saph. Subst. Price & \text{USD} & 40 & +12.5,-12.5 & 10 & +100,-20 & 3 & +66.6,-33.3 & \cref{fig:wafers} \\
        \hline
            Phosphor Prices & \text{USD}/g & 100\% & +50,-50 & 150 & +0,-0 & 200 & +0,-0 & I, \cite{yole_phosphor_2012}\cite{yole2017phosphor} \\
        \hline
        \end{tabularx}
    \vspace{5mm}

    Note: The units for the values in columns \textit{2003}-\textit{2020} are indicated in the column \textit{Units}. If values in columns \textit{2003}-\textit{2020} are instead given in \%, this indicates that the parameters were varied by a set percentage from their respective model baselines. Abbreviations: FTE - full-time equivalent; UPH - units per hour; Equip. Discount - equipment discount (sales rebate for large purchases); Inspec. Yield Savings - yield savings from inspection (early detection and alleviation of issues in the manufacturing workflow).
    \label{tab:sensitivity}
\end{table}

@@

\begin{table}[H]
    \caption{LED innovations and technology improvements affecting key white LED device sub-efficiencies. The table includes innovations and technology improvements affecting forward voltage efficiency, light extraction efficiency, internal quantum efficiency, and light conversion efficiency. For a list of innovations affecting spectral efficiency, see Table 1 in the main article.}
    \vspace{5mm}
    \begin{tabularx}{\textwidth}{ |l|X|X|l|l| }
        \multicolumn{5}{c}{Forward Voltage Efficiency (VFE) $\eta_{Vf}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of Improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1999 & Indium Tin Oxide \newline Current spreading layer & Contact resistance & Yes & \cite{margalith1999indium}\\
        \hline
            1998 & Digitated electrodes & Contact resistance & No & \cite{steigerwald2001electrode} \\
        \hline
            Ongoing & Epitaxy improvements \newline and better doping & Polarization and \newline bulk resistance & No & I \\
        \hline
            1998 & Silver p-Contacts & Contact resistance & No & \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Light extraction efficiency (LEE) $\eta_{LE}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            $<$ 2003 & Optimization for \newline cavity effects & Reduces self-interference of quantum well & No & I, \cite{Shen2003} \\
        \hline
            1993 & Chip surface randomization & Total reflection and absorption & No & I, \cite{bergh1973surface}\cite{Schnitzer1993} \\
        \hline
            1993 & Thin-film chip architecture & Absorption & No & I, \cite{Schnitzer1993} \\
        \hline
            1996 & Patterned sapphire substrate  & Total reflection and absorption & Yes & I, \cite{Tadatomo2001} \\
        \hline
            Ongoing & Chip design for high LEE & Total reflection and absorption & No & I, \cite{Haerle2004} \\
        \hline
            $\sim$2000 & Silver p-contacts & Absorption & No & I, \cite{kondoh2001nitride} \\
        \hline
        \multicolumn{5}{c}{Internal quantum efficiency (IQE) $\eta_{IQ}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            1994 & Double heterostructure & Higher Radiative recombination probability & No & \cite{Nakamura1994} \\
        \hline
            1996 & Multiple quantum well & Higher Radiative recombination probability & No & \cite{Koike1996} \\
        \hline
            Ongoing & Active region \newline Doping & Radiative recombination probability & No & I, \cite{schubert2018light} \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Radiative recombination probability & No & I \\
        \hline
            Ongoing & Chip architecture \newline Improvements & Radiative recombination probability & No & I \\
        \hline
        \multicolumn{5}{c}{Light Conversion efficiency (CE) $\eta_{C}$} \\
        \hline
            \textit{Year} & \textit{Innovation} & \textit{Area of improvement} & \textit{Spillover} & \textit{Source} \\
        \hline
            Ongoing & Lower current density & Current density & No & I \\
        \hline
            Ongoing & Epitaxy \newline Improvements & Charge distribution \newline Defect density & No & I, \cite{bhardwaj2016progress} \\
        \hline
            Ongoing & Chip architecture \newline improvements  & Charge distribution \newline Defect density & No & I, \cite{Wildeson2017} \\
        \hline
            $<$ 2017 & Defect getting \newline underlayer & Defect Density & No & I, \cite{haller2017burying} \\
        \hline
        \end{tabularx}
    \vspace{5mm}

     Note: \textit{Year} column indicates the first instance of application of corresponding invention in white LEDs. 'Ongoing' indicates improvements that are incremental in nature and have been ongoing since the earliest days of LED manufacturing, with no individual breakthroughs identified. \textit{Spillover} column indicates if the innovation involved technology spillovers. \textit{Source} column indicates the source of information about the innovation or improvement, with 'I' indicating expert interviews as such a source.
    \label{tab:innovations}
\end{table}

@@

\begin{table}[h!]
\small
    \centering
    \caption{Anonymized list of LED experts interviewed for this study. Abbreviations: sr. - senior}
    \vspace{5mm}
    \begin{tabular}{|l|l|l|l|l|}
    \hline
        \textit{\#} & \textit{Sector} & \textit{Role} & \textit{Country} & \textit{Expertise} \\ \hline
        1 & Academia & Sr. researcher & UK & Epitaxy \\ \hline
        2 & Industry & Consultant, former sr. researcher & USA & Device architecture \\ \hline
        3 & Industry & Consultant, former head of R\&D & Germany & Epitaxy \\ \hline
        4 & Academia & Professor & Austria & Phosphors \\ \hline
        5 & Industry & Consultant, former head of R\&D & USA & Device architecture \\ \hline
        6 & consulting & Consultant, former sr. technical advisor & USA & Device architecture \\ \hline
        7 & Academia & Professor & Germany & Phosphors \\ \hline
        8 & government & R\&D manager & USA & Device architecture \\ \hline
        9 & consulting & Consultant & USA & Device applications \\ \hline
        10 & Academia & Professor & France & Device physics \\ \hline
        11 & Industry & Sr scientist, former head of R\&D & USA & Device architecture \\ \hline
        12 & Industry & Principal scientist & Germany & Phosphors \\ \hline
        13 & Industry & Former head of R\&D & USA & Phosphors \\ \hline
    \end{tabular}
    \label{tab:interviews}
\end{table}
